---
title: "NOAA Storm Data Analysis"
Synopsis: I conducted an anlysis on NOAA storm data. I found that natural events related to land likely cause property damage. Storm and extream temparature are related to crop damage. They are also related to fatalities as well. High temparature such as Heat wave migh also cause injuries in addition to those events.
author: "Koji"
date: "2021/11/15"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(gridExtra)
knitr::opts_chunk$set(echo = TRUE)
```

download files and read it
```{r, cache=TRUE}
Sys.setlocale(category = "LC_TIME", "C")
Sys.setenv("LANGUAGE"="e")
path <- getwd()
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
download.file(url, file.path(path, "NOAA_data.csv.bz2"))
NOAA_data <- read.csv("NOAA_data.csv.bz2")
names(NOAA_data)
```
## Data Processing
Across the US, which types of events are most harmful with respect to population health?
calculate mean value for each event type by aggregate function
```{r, cache=TRUE}
fatalities_byevent <- aggregate(FATALITIES ~ EVTYPE, data = NOAA_data, mean, na.rm = TRUE)
injuries_byevent <- aggregate(INJURIES ~ EVTYPE, data = NOAA_data, mean, na.rm = TRUE)

```

Across the US, which types of events have the greatest economic consequences?
convert unit k and m into actual numbers
```{r, cache=TRUE}
# mutate new column reflecting units thousand K and million M
NOAA_data <- mutate(NOAA_data, PROPDMGNM = PROPDMG)
NOAA_data <- mutate(NOAA_data, CROPDMGNM = CROPDMG)
for (i in 1:length(NOAA_data)){
  if (NOAA_data$PROPDMGEXP[i] == "K"){
    NOAA_data$PROPDMGNM[i] <- NOAA_data$PROPDMG[i] * 1000
    if (NOAA_data$PROPDMGEXP[i] == "M"){
      NOAA_data$PROPDMGNM[i] <- NOAA_data$PROPDMG[i] * 1000000
    }
  }
 if (NOAA_data$CROPDMGEXP[i] == "K"){
    NOAA_data$CROPDMGNM[i] <- NOAA_data$CROPDMG[i] * 1000
    if (NOAA_data$CROPDMGEXP[i] == "M"){
      NOAA_data$CROPDMGNM[i] <- NOAA_data$CROPDMG[i] * 1000000
    }
 }
}
```

calculate mean value for each event type by aggregate function
```{r, cache=TRUE}
propdmg_byevent <- aggregate(PROPDMGNM ~ EVTYPE, data = NOAA_data, mean, na.rm = TRUE)
cropdmg_byevent <- aggregate(CROPDMGNM ~ EVTYPE, data = NOAA_data, mean, na.rm = TRUE)
```


## Results
Across the US, which types of events are most harmful with respect to population health?
```{r, cache=TRUE}
max(fatalities_byevent$FATALITIES)
max(injuries_byevent$INJURIES)
fatalities_byevent$EVTYPE[which.max(fatalities_byevent$FATALITIES)]
injuries_byevent$EVTYPE[which.max(injuries_byevent$INJURIES)]
```

find which types of event have the greatest economic consequences
```{r, cache=TRUE}
propdmg_byevent$EVTYPE[which.max(propdmg_byevent$PROPDMGNM)]
cropdmg_byevent$EVTYPE[which.max(cropdmg_byevent$CROPDMGNM)]
```

#What are top 10 harmful event with respect to health and economics
```{r, cache=TRUE}
fatalities_byevent_desc <- arrange(fatalities_byevent, desc(FATALITIES))
injuries_byevent_desc <- arrange(injuries_byevent, desc(INJURIES))
propdmg_byevent_desc <- arrange(propdmg_byevent, desc(PROPDMGNM))
cropdmg_byevent_desc <- arrange(cropdmg_byevent, desc(CROPDMGNM))

```

#Top10 events caused highest property damage
```{r, cache=TRUE}
top10_events_prop <- head(propdmg_byevent_desc, 10)

p1 <- ggplot(top10_events_prop, aes(reorder(EVTYPE,PROPDMGNM),PROPDMGNM))+ geom_col(aes(fill = PROPDMGNM)) + scale_fill_gradient2(low = "white", high = "blue") + coord_flip() + theme_classic() + labs(y = "Property damage", x = "Event Type") + ggtitle("Top10 events caused highest PROPERTY damage")
```

#Top10 events caused highest crop damage

```{r, cache=TRUE}
top10_events_crops <- head(arrange(cropdmg_byevent,desc(CROPDMGNM)), 10)
p2 <- ggplot(top10_events_crops, aes(reorder(EVTYPE,CROPDMGNM),CROPDMGNM))+ geom_col(aes(fill = CROPDMGNM)) + scale_fill_gradient2(low = "white", high = "red") + coord_flip() + theme_classic() + labs(y = "Crop damage", x = "Event Type") + ggtitle("Top10 events caused highest CROP damage")
grid.arrange(p1, p2, ncol = 1)
```

#Top10 harmful events with fatalities

```{r, cache=TRUE}
top10_events_fatalities <- head(arrange(fatalities_byevent, desc(FATALITIES)), 10)
p3 <- ggplot(top10_events_fatalities, aes(reorder(EVTYPE,FATALITIES),FATALITIES))+ geom_col(aes(fill = FATALITIES)) + scale_fill_gradient2(low = "white", high = "black") + coord_flip() + theme_classic() + labs(y = "FATALITIES", x = "Event Type") + ggtitle("Top10 harmful events with FATALITIES")
```
#Top10 harmful events with INJURIES

```{r, cache=TRUE}
top10_events_injuries <- head(arrange(injuries_byevent, desc(INJURIES)), 10)
p4 <- ggplot(top10_events_injuries, aes(reorder(EVTYPE,INJURIES),INJURIES))+ geom_col(aes(fill = INJURIES)) + scale_fill_gradient2(low = "white", high = "purple") + coord_flip() + theme_classic() + labs(y = "INJURIES", x = "Event Type") + ggtitle("Top10 harmful events with INJURIES")
grid.arrange(p3, p4, ncol = 1)
```