---
title: "Reproducible research course project week4"
Synopsis: 
author: "Koji"
date: "2021/11/15"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
```

download files and read it
```{r, cache=TRUE}
Sys.setlocale(category = "LC_TIME", "C")
Sys.setenv("LANGUAGE"="e")
path <- getwd()
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
download.file(url, file.path(path, "NOAA_data.csv.bz2"))
NOAA_data <- read.csv("NOAA_data.csv.bz2")
names(NOAA_data)
```
## Data Processing
Across the US, which types of events are most harmful with respect to population health?
calculate mean value for each event type by aggregate function
```{r, cache=TRUE}
fatalities_byevent <- aggregate(FATALITIES ~ EVTYPE, data = NOAA_data, mean, na.rm = TRUE)
injuries_byevent <- aggregate(INJURIES ~ EVTYPE, data = NOAA_data, mean, na.rm = TRUE)

```

Across the US, which types of events have the greatest economic consequences?
convert unit k and m into actual numbers
```{r, cache=TRUE}
# mutate new column reflecting units thousand K and million M
NOAA_data <- mutate(NOAA_data, PROPDMGNM = PROPDMG)
NOAA_data <- mutate(NOAA_data, CROPDMGNM = CROPDMG)
for (i in 1:length(NOAA_data)){
  if (NOAA_data$PROPDMGEXP[i] == "K"){
    NOAA_data$PROPDMGNM[i] <- NOAA_data$PROPDMG[i] * 1000
    if (NOAA_data$PROPDMGEXP[i] == "M"){
      NOAA_data$PROPDMGNM[i] <- NOAA_data$PROPDMG[i] * 1000000
    }
  }
 if (NOAA_data$CROPDMGEXP[i] == "K"){
    NOAA_data$CROPDMGNM[i] <- NOAA_data$CROPDMG[i] * 1000
    if (NOAA_data$CROPDMGEXP[i] == "M"){
      NOAA_data$CROPDMGNM[i] <- NOAA_data$CROPDMG[i] * 1000000
    }
 }
}
head(NOAA_data)
```

calculate mean value for each event type by aggregate function
```{r, cache=TRUE}
propdmg_byevent <- aggregate(PROPDMGNM ~ EVTYPE, data = NOAA_data, mean, na.rm = TRUE)
cropdmg_byevent <- aggregate(CROPDMGNM ~ EVTYPE, data = NOAA_data, mean, na.rm = TRUE)
summary(propdmg_byevent)
summary(cropdmg_byevent)
```


## Results
Across the US, which types of events are most harmful with respect to population health?
```{r, cache=TRUE}
summary(fatalities_byevent)
summary(injuries_byevent)
max(fatalities_byevent$FATALITIES)
max(injuries_byevent$INJURIES)
fatalities_byevent$EVTYPE[which.max(fatalities_byevent$FATALITIES)]
injuries_byevent$EVTYPE[which.max(injuries_byevent$INJURIES)]
```

find which types of event have the greatest economic consequences
```{r, cache=TRUE}
propdmg_byevent$EVTYPE[which.max(propdmg_byevent$PROPDMGNM)]
cropdmg_byevent$EVTYPE[which.max(cropdmg_byevent$CROPDMGNM)]
```

##What are top 5 harmful event with respect to health and economics
```{r, cache=TRUE}
fatalities_byevent_desc <- arrange(fatalities_byevent, desc(FATALITIES))
injuries_byevent_desc <- arrange(injuries_byevent, desc(INJURIES))
propdmg_byevent_desc <- arrange(propdmg_byevent, desc(PROPDMGNM))
cropdmg_byevent_desc <- arrange(cropdmg_byevent, desc(CROPDMGNM))

```

#Top10 events caused highest property damage
```{r, cache=TRUE}
top10_events_prop <- head(propdmg_byevent_desc, 10)

ggplot(top10_events_prop, aes(reorder(EVTYPE,PROPDMGNM),PROPDMGNM))+ geom_col(aes(fill = PROPDMGNM)) + scale_fill_gradient2(low = "white", high = "blue") + coord_flip() + theme_classic() + labs(y = "Property damage", x = "Event Type") + ggtitle("Top10 events caused highest property damage")
```

#Top10 events caused highest crop damage

```{r, cache=TRUE}
top10_events_crops <- head(arrange(cropdmg_byevent,desc(CROPDMGNM)), 10)
ggplot(top10_events_crops, aes(reorder(EVTYPE,CROPDMGNM),CROPDMGNM))+ geom_col(aes(fill = CROPDMGNM)) + scale_fill_gradient2(low = "white", high = "red") + coord_flip() + theme_classic() + labs(y = "Crop damage", x = "Event Type") + ggtitle("Top10 events caused highest crop damage")
```

#Top10 harmful events with fatalities

```{r, cache=TRUE}
top10_events_fatalities <- head(arrange(fatalities_byevent, desc(FATALITIES)), 10)
ggplot(top10_events_fatalities, aes(reorder(EVTYPE,FATALITIES),FATALITIES))+ geom_col(aes(fill = FATALITIES)) + scale_fill_gradient2(low = "white", high = "black") + coord_flip() + theme_classic() + labs(y = "FATALITIES", x = "Event Type") + ggtitle("Top10 harmful events with fatalities")
```
#Top10 harmful events with INJURIES

```{r, cache=TRUE}
top10_events_injuries <- head(arrange(injuries_byevent, desc(INJURIES)), 10)
ggplot(top10_events_injuries, aes(reorder(EVTYPE,INJURIES),INJURIES))+ geom_col(aes(fill = INJURIES)) + scale_fill_gradient2(low = "white", high = "purple") + coord_flip() + theme_classic() + labs(y = "INJURIES", x = "Event Type") + ggtitle("Top10 harmful events with INJURIES")
```